\section{Theory}

Responsible for the risk of instability of control circuits are the delays. It
is therefore always necessary to ensure that a  system  is  stable. To achieve
this we need to tune a PID controller.

Tuning  a PID controller  means  to  find  suitable  parameters  for  the  PID
controller  for  a  given  application  so  that the application can  be  very
effective. This is achieved by using Ziegler-Nichols and Chien-Hrones-Reswick.


\subsection{Tuning Rules According to Ziegler-Nichols (Oscillation Method)}

In this method it  is  assumed  that  there  are  higher-orders,  which can be
described with a series connection of dead time element $T_{t}$ and a $PT_{1}$
element.

\begin{equation}
    G_{s}(s) =  \approx \frac{K_{PS}}{1+sT_{1}} \cdot e^{-sT_{1}}
\end{equation}

If the controller  is  tentative  operated  with  a  P-controller,  so one can
determine  the  critical  $K_{P}$  and  the  critical period $\tau_{crit}$  by
increasing $K_{P}$ until the stability limit is reached.
The procedure of Ziegler-Nichols is:

\begin{enumerate}
    \item Use a pure P controller and set a small gain $K_{P}$.
    \item Increasing the gain $K_{P}$ until an undamped oscillation occurs.
    \item Indentify the critical gain $K_{P,crit}$ and the critical period $\tau_{crit}$
    \item According to table \ref{tab:ziegler_nichols} specifying the controllers parameter.
\end{enumerate}

\begin{center}
    \begin{threeparttable}
        \begin{tabular}{cccc}
            \toprule
            Controller type & $K_{P}$ & $T_{i}$ & $T_{d}$ \\
            \midrule
            P    &  $0.5  \cdot K_{P,crit}$  &  -                         &  -                         \\
            PI   &  $0.45 \cdot K_{P,crit}$  &  $0.85 \cdot \tau_{crit}$  &  -                         \\
            PID  &  $0.6  \cdot K_{P,crit}$  &  $0.5 \cdot \tau_{crit}$   &  $0.12 \cdot \tau_{crit}$  \\
            \bottomrule
        \end{tabular}
        \caption{Table with controller parameters}
        \label{tab:ziegler_nichols}
    \end{threeparttable}
\end{center}

The advantage of the Ziegler-Nichols is that he is simple to apply and because
of that also  easy  to  understand. On the other side it is very risky because
the control loop must be operated close to instability.  This can be depending
on the system very dangerous and very expensive. Furthermore theoretically not
all control plants can be made to oscillate  even  if  in  the  practice  most
plants  oscillate  because  of  delays  and  non  linearities.  The  practical
significance of this method is thus limited.


\subsection{Tuning Rules According to Chien-Hrones-Reswick (Step Response Method)}

The step response method is  an  improvement  of  the  Ziegler-Nichols method.
Before using the step  response  method,  first  we  have  to  select the step
amplitude. If the step amplitude is too small, the control plant behaves in an
atypic way. This happens because of  small  disturbances which affect the step
response. A good example for this is static friction. On the other side if the
step amplitude is too large, then nonlinear effects can affect the behavior of
the control plant.

The procedure of Chien-Hrones-Reswick is:

\begin{enumerate}
    \item Recording the step response of the control plan and measure the gain $K_{S}$, the delay time $T_{u}$ and the compensation time $T_{g}$.
    \item Determine the controller parameters according to the tuning rule (see \ref{tab:chien_hrones_reswick}
\end{enumerate}

\begin{center}
    \begin{threeparttable}
        \begin{tabular}{cccc}
            \toprule
            Controller type & $K_{P}$ & $T_{i}$ & $T_{d}$ \\
            \midrule
            P    &  $0.3  \cdot\frac{T_{g}}{T_{u}} \cdot \frac{1}{K_{s}}$  &  -                  &  -                  \\
            PI   &  $0.35 \cdot\frac{T_{g}}{T_{u}} \cdot \frac{1}{K_{s}}$  &  $1.2 \cdot T_{g}$  &  -                  \\
            PID  &  $0.3  \cdot\frac{T_{g}}{T_{u}} \cdot \frac{1}{K_{s}}$  &  $T_{g}$            &  $0.5 \cdot T_{u}$  \\
            \bottomrule
        \end{tabular}
        \caption{Table with Controller Parameters(Aperiodic behavior, good tracking)}
        \label{tab:chien_hrones_reswick}
    \end{threeparttable}
\end{center}

Based on the ratio of $\frac{T_{u}}{T_{g}}$,  we  can  say how controllability
the control plant is.

\begin{center}
    \begin{threeparttable}
        \begin{tabular}{ccc}
            \toprule
            $\frac{T_{u}}{T_{g}}$                 &  Controllability  & Effort \\
            $\frac{T_{u}}{T_{g}} < 0.1$           &  very good        & small  \\ 
            $0.1 \leq \frac{T_{u}}{T_{g}} < 0.2$  &  good             & medium  \\ 
            $0.2 \leq \frac{T_{u}}{T_{g}} < 0.4$  &  fair             & large  \\ 
            $0.4 \leq \frac{T_{u}}{T_{g}} < 0.8$  &  bad              & very large  \\ 
            $ \frac{T_{u}}{T_{g}} \geq 0.2$       &  very bad         & special measures required  \\ 
        \end{tabular}
        \caption{Table with Controller Parameters(Aperiodic behavior, good tracking)}
        \label{tab:controllability}
    \end{threeparttable}
\end{center}

